---
title: "BioTime"
author: "Kari Norman"
date: "4/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(lazyeval)
library(rdataretriever)

```

Import occurence data from Biotime
```{r}
#rdataretriever::install("biotime", connection = "csv", data_dir = here("data"))

data <- read_csv(here("data", "biotime_metadata.csv")) %>% filter(taxa %in% c("Birds", "Mammals")) %>%
  select(study_id, taxa) %>%
  left_join(read_csv(here("data", "biotime_query.csv")), by = "study_id")
  
```

Get mammal and trait data
```{r}
bird_trait <- read_csv(here("data/elton_traits/elton_traits_BirdFuncDat.csv"))
mamm_trait <- read.csv(here("data/elton_traits/elton_traits_MammFuncDat.csv"))
```

Check which authority would result in the most matches 
```{r}
library(taxadb)

get_match_counts <- function(data, species_col){
  species_col <- lazyeval::as_name(species_col)
  data %>% 
    select(!!species_col) %>% 
    drop_na() %>%
    distinct() %>%
    mutate(
      gbif = get_ids(!!species_col, "gbif"),
      col = get_ids(!!species_col, "col" ),
      itis = get_ids(!!species_col, "itis"),
      ncbi = get_ids(!!species_col, "ncbi"),
      wd = get_ids(!!species_col, "wd"  ),
      iucn = get_ids(!!species_col, "iucn"),
      ott = get_ids(!!species_col, "ott" ) 
    ) %>%
    #select(!!call("-", species_col)) %>%
    purrr::map_dbl(function(x) sum(!is.na(x)))
}

##Counts show that OTT is the best authority for both data sets
get_match_counts(data %>% filter(taxa == "Birds"), "genus_species")
get_match_counts(data, "genus_species")
get_match_counts(bird_trait, "scientific")
get_match_counts(mamm_trait, "scientific")
```

OTT is the best authority but doesn't have common names, so we use the next best ITIS. We'll start with BioTime data:

##Biotime

```{r}
#Getting matches by scientific name 
biotime <- data %>% 
  #filter(taxa == "Birds") %>%
  select(genus_species) %>%
  drop_na() %>%
  distinct() %>%
  mutate(id = get_ids(genus_species, "itis")) %>%
  left_join(data, by = "genus_species")

#Getting matches by sci name and common name 
biotime_ids <- by_name(unique(data$genus_species), "itis") %>%
  bind_rows(by_common(unique(data$genus_species), "itis")) %>%
  filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Some names match to multiple acceptedUsage ID's, so we have to manually choose which one we want. In this case it's just the Blue-winged warbler
```{r}

get_dupe_ids <- function(data){
  biotime_ids %>%
    distinct(acceptedNameUsageID, input, sort) %>%
    group_by(sort) %>% 
    filter(n()>1) %>% 
    ungroup()
}  

unres <- get_dupe_ids(biotime_ids)
unres 
```

Add the unresolved ID to all the other ids and join with biotime data
```{r}
biotime_ids <- unres %>%
left_join(biotime_ids) %>% #join ID's back with all the columns
  filter(taxonomicStatus == "accepted") %>% #want the accepted ID from the two Warbler options
  bind_rows(biotime_ids %>% filter(!sort %in% unres$sort)) #join to resolved ID's, excluding the observations we resolved manually

biotime_data <- biotime_ids %>%
  select(id = acceptedNameUsageID,input) %>%
  distinct() %>%
  right_join(data, by = c("input" = "genus_species")) %>%
  rename(genus_species = input)
```



##Elton Traits
Get id's for the Elton Bird trait data
```{r}
bird_ids <- by_name(unique(bird_trait$scientific), "itis") %>%
  bind_rows(by_common(unique(bird_trait$scientific), "itis")) %>%
  #filter(taxonRank == "species") %>% #only want ID's to species, since that's the level of the trait data
  drop_na(acceptedNameUsageID)
```

Are there any unresolved ids? Yes - just one, which is a synonym to two different accepted id's
```{r}
unres_trait <- get_dupe_ids(bird_ids) %>%
  left_join(bird_ids)

unres_trait

#resolve to the desired ID (in this case the one with complete hierarchy data)
bird_ids <- unres_trait %>% 
  filter(!is.na(kingdom)) %>%
  bind_rows(bird_ids %>% filter(!sort %in% unres_trait$sort))
  
  
bird_trait_data <- bird_ids  %>%
  select(id = acceptedNameUsageID, input) %>%
  distinct() %>%
  right_join(bird_trait, by = c("input" = "scientific")) %>%
  rename(scientific = input)
```

check if all bird biotime data w/id's has an elton match. There are 54 species without a match
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id) %>%
  distinct() %>%
  filter(!id %in% bird_trait_ids$acceptedNameUsageID) %>%
  dim()
```

and 605 species without an ID
```{r}
biotime_data %>%
  filter(taxa == "Birds") %>%
  select(id, genus_species) %>%
  distinct() %>%
  filter(is.na(id)) %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
bird_data <- biotime_data %>%
  filter(taxa == "Birds") %>%
  rowid_to_column("row") %>%
  left_join(bird_trait_data, na_matches = "never")
```

