---
title: "fishtraits"
author: "Kari Norman"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RSelenium)
```

```{r}
clickObject <- function(selector){
  object <- rd$findElement(using = 'css selector', selector)
  object$clickElement()
}
```
connect to Selenium server and navigate to the page
```{r}
driver <- rsDriver(browser = c("chrome"), chromever = "76.0.3809.126")
rd <- driver[["client"]]
rd$open()

rd$navigate("http://www.fishtraits.info/search/taxo/")
```

see if we can get the list of all possible search items for the FID dropdown (second is dependent on the first input)
```{r}
#click on drop down element
clickObject('.ui-combobox .ui-icon-triangle-1-s')

#get text of the list
fid_list <- rd$findElement('css selector', '#ui-id-1')
fid_text <- unlist(strsplit(unlist(fid_list$getElementText()), "\n"))
```

Choose FID and Genus. All the boxes under the "columns" header unfortunately need to be checked manually because css selectors couldn't be identified for each individual box. Only have to do it once though!
```{r}

for (i in 1:length(fid_text)) {
  #select and search the FID dropdown
  fid <- rd$findElement(using = 'css selector', "#fid-combo")
  fid$clearElement()
  #pass to search box
  print(fid_text[i])
  fid$sendKeysToElement(list(fid_text[i]))
  #click matching one
  Sys.sleep(0.25)
  clickObject("#ui-id-1")
  
  #get list of Genus's that are now possible
  clickObject(".ui-combobox2 .ui-icon-triangle-1-s")
  Sys.sleep(0.25)
  genus_list <- rd$findElement('css selector', '#ui-id-2')
  genus_text <- unlist(strsplit(unlist(genus_list$getElementText()), "\n"))
  print(genus_text)
  
  for (j in 1:length(genus_text)) {
    #select and search the Genus dropdown
    genus <- rd$findElement(using = 'css selector', "#genus-combo")
    genus$clearElement()
    genus$sendKeysToElement(list(genus_text[j]))
    Sys.sleep(0.15)
    #click matching one
    clickObject('#ui-id-2')

    #select search box
    clickObject("#searchSubmit")
    #and download
    clickObject("#downloadbtn")
  }
}
```

```{r}
driver$server$stop()
```

Move all the files to a folder in the appropriate place
```{r}
library(filesstrings)
file_names <- list.files("~/Downloads", "fishtraits*")

dir.create(here::here("inst", "extdata", "fishtraits"))
file.move(paste0("~/Downloads/", file_names), here::here("inst", "extdata", "fishtraits"))
```

