---
title: "Merge biotime timeseries with trait data"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Description
This script merges trait data and biotime timeseries data, then filters out timeseries that don't have high enough trait coverage. 

```{r}
library(tidyverse)
library(FD)
devtools::load_all()
```

```{r}
pins::board_register_github(name = "github", repo = "karinorman/biodivTS_data", branch = "master")

biotime_data <- pins::pin_get("biotime-data", board = "github")
elton_mamm <- pins::pin_get("elton-mamm", board = "github")
elton_bird <- pins::pin_get("elton-bird", board = "github")
amphibio <- pins::pin_get("amphibio", board = "github")

biotime_data <- biotime_data %>%
  ungroup() %>%
  mutate(obsID = row_number())
```

check if all bird & mammal biotime data w/id's has an elton match. There are 13 species without a match.
```{r}
no_trait <- biotime_data %>%
  filter(taxa %in% c("Birds", "Mammals")) %>%
  select(id, sourceName, taxa) %>%
  drop_na(id) %>%
  distinct() %>%
  filter(!id %in% elton_bird$id, !id %in% elton_mamm$id)

dim(no_trait)
```

Function for joining timeseries data with functional trait data. 
```{r}
merge_data <- function(trait_data){
  biotime_data %>%
    select(-scientificName) %>% #remove scientific name from biotime data
    rename(biotimeName = sourceName) %>% #give original biotime name a better column name
  left_join(trait_data %>% rename(traitName = sourceName), by = "id",
            na_matches = "never") %>% # join to trait data
   drop_na(scientificName)
}
```

Get data w/traits for all studies that have >75% trait coverage

`id` is `NA` if there was no `id` match
`biotimeName` or `traitName` is `NA` if that ID is the result of an average of multiple species matches
`scientificName` is `NA` when there was no trait data for a given ID
```{r}
#set coverage threshold for analysis dataset
percent <- 0.75

traits <- list(elton_bird, elton_mamm, amphibio)

#map across list of trait sources to merge timeseries with each one
trait_merge <- map_dfr(traits, merge_data) 

#combine observations w/trait data back with observations with no trait data
comp_biotime_traits <- trait_merge %>%
  bind_rows(biotime_data %>% 
              filter(!obsID %in% trait_merge$obsID) %>%
              select(-scientificName) %>%
              rename(biotimeName = sourceName))# %>%
  #distinct()

#always check that one species didn't match to multiple datasets, below should return and empty dataframe
#comp_biotime_traits %>% group_by(record_id) %>% filter(n() > 1) %>% dim()

#get a percentage of trait coverage for each study
coverage <- comp_biotime_traits %>%
  select(study_id, id, biotimeName, traitName) %>%
  distinct() %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(traitName))/n()) %>%
  arrange(desc(perc))

#studies to include, since they meet the coverage requirement 
inc_coverage <- coverage %>% filter(perc > percent)

biotime_traits <- comp_biotime_traits %>%
  #get studies with appropriate coverage and drop observations that we don't have trait data for, and unlabeled non-id'd species
  filter(study_id %in% inc_coverage$study_id, !is.na(scientificName)) %>% 
  group_by(rarefyid, year) %>%
  filter(n_distinct(id) > 4) %>% #more than four species for each observation
  ungroup()
```

Get final columns 
```{r}
bt_cols <- colnames(biotime_data)[!colnames(biotime_data) %in% 
                                         c("sourceName", "obsID", "scientificName")]
biotime_traits <- biotime_traits %>% 
  select(all_of(bt_cols)) %>%
  rename(species = id)
```

Some timeseries have observations from more than one class of species (e.g. birds and mammals), here we split them into two individual timeseries.
```{r}
#get the class of each species
tax_info <- taxadb::filter_id(unique(biotime_traits$species), "itis")
biotime_traits <- biotime_traits %>% 
  left_join(tax_info %>% select(input, class), by = c("species" = "input"))

#find all the studies with more than one taxa type
mix_study <- biotime_traits %>% 
  select(species, class, taxa, study_id) %>% 
  distinct() %>% 
  count(study_id, class, taxa) %>% 
  group_by(study_id) %>%
  filter(n_distinct(class, na.rm = TRUE) > 1) %>% 
  pull(study_id) %>%
  unique()

#split into two different timeseries based on whether they're bird or mammal observations,
#remove years with too few species
mix_split <- biotime_traits %>%
  filter(study_id %in% mix_study) %>%
  mutate(rarefyid = if_else(class == "Aves", paste0(rarefyid, "_bird"), paste0(rarefyid, "_mamm")),
         taxa = if_else(class == "Aves", "Birds", paste0(rarefyid, "Mammals"))) %>%
  group_by(rarefyid, year) %>% 
  filter(n_distinct(species) > 4)

#update the studies that needed to be split with new rarefyid breakdowns
bt_traitfiltered <- biotime_traits %>%
  filter(!study_id %in% mix_study) %>%
  bind_rows(mix_split) %>%
  select(-class)
```

Final merged data set to save
```{r}
#save locally and in cache
usethis::use_data(bt_traitfiltered)
pins::pin(bt_traitfiltered, board = "github")
```

Create a reference table of trait data
```{r}
trait_cols <- c(colnames(amphibio), colnames(elton_bird), colnames(elton_mamm))
trait_cols <- trait_cols[!trait_cols %in% c("sourceName", "scientificName")]

trait_ref <- biotime_traits %>%
  select(all_of(trait_cols)) %>%
  rename(Species = id) %>%
  distinct()

#save locally and in cache
usethis::use_data(trait_ref)
pins::pin(trait_ref, board = "github")
```

Create table of info for each study
```{r}
metadata <- pins::pin_get("metadata", board = "github")

study_table <- biotime_traits %>%
  select(study_id, plot) %>%
  count(study_id, name = "n_plots") %>% 
  left_join(metadata %>% select(study_id, realm, climate, habitat, biome_map, taxa, 
                                organisms, start_year, end_year), by = "study_id") %>%
  left_join(coverage %>% select(study_id, coverage = perc), by = "study_id")

#save locally and in cache
usethis::use_data(study_table)
pins::pin(study_table, board = "github")
```
