---
title: "Check Coverage"
author: "Kari Norman"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(FD)
```

```{r}
load(here::here("data", "biotime_data.rda"))
load(here::here("data", "elton_mamm.rda"))
load(here::here("data", "elton_bird.rda"))
load(here::here("data", "amphibio.rda"))
```

#Checking Coverage of ID's and trait data -> biotime

check if all bird & mammal biotime data w/id's has an elton match. There are 63 species without a match (54 are birds).
```{r}
no_trait <- biotime_data %>%
  filter(taxa %in% c("Birds", "Mammals")) %>%
  select(id, sourceName, taxa) %>%
  drop_na(id) %>%
  distinct() %>%
  filter(!id %in% elton_bird$id, !id %in% elton_mamm$id)

dim(no_trait)
```

and 19869 species without an ID, some of which are species codes
```{r}
biotime_data %>%
  #filter(taxa == "Birds") %>%
  select(id, sourceName, taxa) %>%
  filter(is.na(id)) %>%
  distinct() %>%
  dim()
```

Are there any studies for which we have trait data for all species?
```{r}
merge_data <- function(taxa, trait_data){
  biotime_data %>%
    select(-sourceName, -scientificName) %>%
  filter(taxa == taxa) %>%
  left_join(trait_data, by = "id",
            na_matches = "never")
}

get_coverage <- function(taxa_data){
  coverage <- taxa_data %>%
  group_by(study_id) %>%
  summarise(perc = sum(!is.na(sourceName))/n()) %>%
  arrange(desc(perc))
}

get_coverage(merge_data("Birds", elton_bird)) %>% View
get_coverage(merge_data("Mammals", elton_mamm)) %>% View()
get_coverage(merge_data("Amphibians", amphibio)) %>% View()
```

Get data w/traits for all studies that have >85% trait coverage
```{r}
#set coverage threshold for analysis dataset
percent <- 0.85

traits <- list(elton_bird, elton_mamm, amphibio)
taxa <- list("Birds", "Mammals", "Amphibians")

biotime_traits_list <- map2(taxa, traits, ~merge_data(taxa = .x, trait_data = .y))
coverage <- map_dfr(biotime_traits_list, get_coverage) %>%
  filter(perc > percent)

biotime_traits <- map_dfr(biotime_traits_list, ~drop_na(., scientificName)) %>%
  filter(study_id %in% coverage$study_id)
```

```{r}
usethis::use_data(biotime_traits)
```

Get FD metrics for each study 
```{r}
load(here::here("data", "biotime_traits.rda"))

get_fd <- function(data, study, plot){
  print(max(data$record_id))
  species_mat <- data %>%
    select(year, id) %>% 
    distinct() %>%
    arrange(year, id) %>%
    mutate(present = 1) %>%
    spread(id, value = present, fill = 0) %>%
    select(-year)
  
  years <- data %>% 
    select(year) %>%
    distinct() %>%
    arrange(year)
  
  trait_mat <- data %>%
    select(-c(taxa, record_id, day, month, year, sample_desc, id_species, 
              latitude, longitude, sum_abundance, sum_biomass, sourceName, scientificName)
    ) %>%
    distinct() %>%
    select_if(~ length(unique(.)) > 1) %>% #remove columns that have the same value for all columns
    mutate_at(vars(-id, -ends_with("5cat")), list(~scales::rescale(.,to=c(0, 1)))) %>% #rescale variables
    arrange(id) %>%
    column_to_rownames("id")
  
  dir.create(paste0(here::here("data"), "/fd_output"), showWarnings = FALSE)
  
  as.tibble(dbFD_joggle(trait_mat, species_mat, w.abun = FALSE))[,1:8] %>%
    cbind(years, .) %>%
    mutate(study_id = study, plot = plot) %>%
    write_tsv(paste0(here::here("data"), "/fd_output/", "s", study, "_p", plot, "_fd.tsv"))
}

get_fd_safe <- possibly(get_fd, NA)

fd_data <- biotime_traits %>% 
  filter(!is.na(-plot)) %>%
  group_by(study_id, year, plot) %>%
  filter(n_distinct(id) > 4) %>%
  ungroup() %>%
  group_by(study_id, plot) %>%
  group_map(~get_fd_safe(.x, study = .y$study_id, plot = .y$plot))
```

Read data back in:
```{r}
fd_data <- dir(here::here("data", "fd_output"), "*.tsv") %>%
  paste0(here::here("data", "fd_output"), "/", .) %>%
  map_dfr(read_tsv)
```


