---
title: "null_model"
author: "Kari Norman"
date: "6/2/2020"
output: html_document
---

```{r}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Create null models that would be transferable to any timeseries by treating the species pool as all species observed across the timeseries.

Simulation functions
```{r}
load(here::here("data", "biotime_traits.rda"))

get_plot_sample <- function(data, study, plot_id, n){
  rich_levels <- data %>%
    filter(study_id == study, plot == plot_id) %>%
    select(year, id) %>%
    distinct() %>%
    count(year) %>%
    pull(n) %>%
    unique()
  
  species_pool <- data %>% 
    filter(study_id == study, plot == plot_id) %>%
    pull(id) %>%
    unique() #get column of unique scientific names
  
  get_samp_species <- function(richness){
    samp_species <- sample(species_pool, richness)
    df <- cbind(richness, samp_species)
    return(df)
  }
  
  site_samples <- data.frame()
  for (j in 1:n){
    iter_samples <- data.frame()
    for(i in rich_levels){
    iter_samples <- rbind(iter_samples, get_samp_species(i))
    }
    site_samples <- rbind(site_samples, iter_samples %>% mutate(n = j))
  }
  site_samples %>%
    mutate(study_id = study, plot = plot_id, value = 1) %>%
    spread(samp_species, value)
 # return(site_samples %>% mutate(study_id = study, plot = plot_id))
}

#drop trait columns that are the same value for every observation, and therefore providing no additional info
clean_trait_matrix <- function(trait_matrix){
  uniq_vals <- sapply(trait_matrix, function(x){length(unique(x))}) #get number of unique values of each trait
  col_names <- names(which(uniq_vals == 1)) #find columns where there is only one value for all observations
  return(select(trait_matrix, -col_names))
}

get_sample_FD <- function(data, n){
  #get random sample for each region and richness level
  plots <- data %>%
    select(site_id, plot) %>%
    distinct()
  
  sample_occurrence <- pmap_dfr(plots[1:10,], get_plot_sample) 
  
  sample_species_mat <- select(sample_occurrence, -richness, -site_id) %>% #want only species x site info
    select(sort(current_vars())) #sort columns to be in alphabetical order, required for dbFDI()
  sample_trait_mat <- clean_trait_matrix(get_trait_matrix(colnames(sample_species_mat)))

  sample_fd <- as.data.frame(dbFD_joggle(x = sample_trait_mat, a = sample_species_mat, w.abun = FALSE))
 
  df <- sample_occurrence %>%
    select(region, richness) %>%
    bind_cols(sample_fd) %>%
    mutate(merge_test = case_when(
      richness == nbsp ~ TRUE,
      TRUE ~ FALSE
    ))
  
  if(sum(df$merge_test) != dim(df)[1]) warning("richness and region columns may not have been appropriately joined with FD data")
  
  path <- paste0(here(), "/data/fd_site_samples/", x, "_tmp.tsv.bz2")
  write_tsv(df, path)
  #pb_upload(path)
}
```

```{r}
n <- 100 #sample size

if(!dir.exists(here::here("data", "fd_site_samples"))){
  dir.create(paste0(here::here(), "/data/fd_site_samples"))
  system.time({out <- future_map(1:n, get_sample_FD)})
}

sim <- purrr::map_dfr(fs::dir_ls(path = here::here("data", "fd_samples"), glob="*.tsv.bz2"), readr::read_tsv, .id = "sample")

```


```

